\chapter*{References}
\label{ch:References}

\begin{description}

\item[Andersson-Sk{\"o}ld and Simpson, 1997] Y. Andersson-Sk\"{o}ld
  and D. Simpson. Comparison of the chemical schemes of the {EMEP MSC-W}
		  and the {IVL} photochemical trajectory models. EMEP
                  MSC-W Note 1/97. The Norwegian
Meteorological Institute, Oslo, Norway, 1997.

\item[Emberson {\sl et al.}, 2000] L. Emberson, D. Simpson,
  J.-P. Tuovinen, M.R. Ashmore and H.M. Cambridge. Towards a model of
  ozone deposition and stomatal uptake over Europe. EMEP/MSC-W
  Note 6/2000.

\item[Fagerli {\sl et al.}, 2004] H. Fagerli, D. Simpson and
  S. Tsyro.
 Unified EMEP model: Updates. In {\it EMEP Status Report
    1/2004, Transboundary acidification, eutrophication and ground
    level ozone in Europe. Status Report 1/2004}, pages 11--18. The Norwegian
Meteorological Institute, Oslo, Norway, 2004.

\item[Gardner {\sl et al.}, 1997] R.M. Gardner, K. Adams, T. Cook,
  F. Deidewig, S. Ernedal, R. Falk, E. Fleuti, E. Herms, C.E. Johnson,
  M. Lecht, D.S. Lee, M. Leech, D. Lister, B. Masse, M. Metcalfe,
  P. Newton, A. Schmitt, C. Vandenbergh and R. Van Drimmelen. The
  ANCAT/EC global inventory of NO$_x$ emissions from 
		  aircraft. {\it Atmos. Environ.}, 31 (12): 1751-1766
, 1997.


\item[Jonson {\sl et al.}, 2001] J.E.Jonson, J.K.Sundet and
  L. Tarras{\'o}n. Model calculations of present and future levels of
ozone and ozone precursors with a global and a regional model. {\it
  Atmos. Environ.}, 35: 525-537, 2001.  


\item[K{\"o}hler {\sl et al.}, 1995] I. K{\"o}hler, R. Sausen and
  G. Klenner. 
NO$_x$ production from lightning. The impact of NO$_x$ emissions from
aircraft upon the atmosphere at flight altitudes 8-15 km (AERONOX),
edited by U. Schumann, final report to the Commission of the European
Communities, Deutch Luft und Raumfart, Oberpfaffenhofen, Germany,
343--345, 1995.


\item[Logan, 1998] J. A. Logan. An analysis of ozonesonde data for the
  troposhere: Recommendations for testing 3-D models and development
  of a gridded climatology for troposheric ozone. {\it
    J. Geophys. Res.}, 10 (D13):16, 115--116, 149, 1998.


\item[Reis {\sl et al.}, 2000] S. Reis, D. Simpson, R. Friedrich,
  J.E. Jonson, S. Unger and A. Obermeier. Road traffic emissions -
  predictions of future  contributions to regional ozone levels in Europe.
{\it Atmos. Environ.}, 34: 4701-4710, 2000.

\item[Simpson {\sl et al.}, 2003] D. Simpson, H. Fagerli, J.E. Jonson, 
                    S. Tsyro, P. Wind and J.-P. Tuovinen.
{\it The EMEP Unified Eulerian Model. Model Description. EMEP MSC-W Report
1/2003}. The Norwegian
Meteorological Institute, Oslo, Norway, 2003.


\item[Simpson and Emberson, 2006] D. Simpson and L. Emberson. Ozone
  fluxes -- Updates.  
% L. Tarras{\'o}n,
%  H. Fagerli,  H. Klein, D. Simpson, A.C. Benedictow, V. Vestreng, E. Rigler,
%  L. Emberson, M. Posch and T. Spranger. 
 In {\it Transboundary Acidification,
  Eutrophication  and Ground Level Ozone in Europe from 1990 to 2004
  in support for the review of the Gothenburg Protocol. EMEP Status
                   Report 1/2006}, pages 63--79. The Norwegian
Meteorological Institute, Oslo, Norway, 2006. 

\item[Tarras\'on {\sl et al.}, 2004] L. Tarras{\'o}n, H. Fagerli, 
              J.E. Jonson, H. Klein, M. van Loon,  
                   D. Simpson, S. Tsyro, V. Vestreng,
                   P. Wind, M. Posch, S. Solberg, T. Spranger,
                   P. Thunis and L. White. Transboundary
                   Acidification, Eutrophication 
                   and Ground Level Ozone in Europe. EMEP Status
                   Report 1/2004. The Norwegian
Meteorological Institute, Oslo, Norway, 2004. 



\end{description}
